---
title: "HW 3"
author: Shihui Zhu sz3029
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set()
```

```{r package, message=FALSE}
# This chunk loads all the packages used in this homework
library(tidyverse)
library(p8105.datasets)
```

# Problem 1

## Load instacart dataset
```{r load_instacart}
data("instacart")
```

## Description of the dataset

The instacart dataset contains `r nrow(instacart)` observations of `r ncol(instacart)` variables. The variables are: ``r names(instacart)``. Some key variables are: 

* ``r names(instacart)[1]``: order identifier, contains `r length(unique(dplyr::pull(instacart, order_id)))` individual orders

* ``r names(instacart)[2]``: product identifier, contains `r length(unique(dplyr::pull(instacart, product_id)))` individual products

* ``r names(instacart)[4]``: 1 if the product has been ordered by this user in the past, 0 otherwise

* ``r names(instacart)[5]``: customer identifier, contains `r length(unique(dplyr::pull(instacart, user_id)))` individual customers

* ``r names(instacart)[7]``: the order sequence number for this user

* ``r names(instacart)[8]``: the day of the week on which the order was placed

* ``r names(instacart)[12]``: aisle identifier, contains `r length(unique(dplyr::pull(instacart, aisle_id)))` aisles

* ``r names(instacart)[13]``: department identifier, contains `r length(unique(dplyr::pull(instacart, department_id)))` departments

`r instacart[1,] %>% knitr::kable()`

An example of the above observation in the dataset: A customer with id 112108 ordered "Bulgarian Yogurt" (product id 49302) from department "dairy eggs," aisle "yogurt". This was his/her 4th order at instacart, and the product was a re-ordered product for him/her, as well as the first order added to cart. The order was placed at Thursday 10am, and it was 9 days after his/her last order. 

## Questions

1. How many aisles are there, and which aisles are the most items ordered from?
```{r unique_aisle}
length(unique(dplyr::pull(instacart, aisle_id)))
```

There are 134 aisles.

```{r most_aisle}
instacart %>%
  group_by(aisle_id) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  slice(1:1)
```

The aisle with id 83, "fresh vegetables," is which the most items ordered from, and it is ordered 150609 times.

2. Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r hist}
instacart %>%
  group_by(aisle_id) %>%
  filter(n() > 10000) %>%
  arrange(aisle_id) %>%
  ggplot(aes(x = aisle_id)) +
  geom_histogram(binwidth = 2) +
  labs(
    title = "Histogram for number of items ordered in each aisle (> 10000 items ordered)",
    x = "Aisle ID",
    y = "Number of items ordered")
```

3. Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r most_pop, message=FALSE}
instacart %>%
  filter(
    aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")
  ) %>%
  group_by(aisle, product_name) %>%
  summarise(count = n()) %>%
  mutate(rank =  dense_rank(desc(count))) %>%
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>%
  knitr::kable(col.names = c("Aisle", "Product Name", "Count", "Rank"))
```

4. Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers 

```{r P_C, message=FALSE}
#week_day <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
instacart %>%
  filter(
    product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")
  ) %>%
  mutate(order_dow = lubridate::wday(order_dow + 1, label = TRUE)) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean = mean(order_hour_of_day), .groups = "drop") %>%
  spread(order_dow, mean) %>%
  knitr::kable(digits = 1, 
               col.names = c("Product", "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday"),
               caption = "The mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week")
```

# Problem 2

## Load BRFSS dataset
```{r clean_brfss}
data("brfss_smart2010")

brfss_smart2010 <- 
  brfss_smart2010 %>% 
  janitor::clean_names() %>%
  # rename variables
  rename("location" = locationdesc, 
         "resp_id" = respid) %>%
  separate(location, into = c("state", "location"), sep = " - ") %>%
  select(-locationabbr) %>%
  # focus on the “Overall Health” topic
  filter(topic == "Overall Health") %>% 
  # include only responses from “Excellent” to “Poor”
  filter(
    response %in% c("Excellent", "Poor", "Very good", "Good", "Fair")
    ) %>%
  mutate(
    response = factor(response, 
                      levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), 
                      ordered = TRUE)) %>%
  # organize responses as a factor taking levels ordered from “Poor” to “Excellent”
  arrange(response)

brfss_smart2010
```

1. In 2002, which states were observed at 7 or more locations? 
```{r 2002, message=FALSE}
more_than_7_2002 <-
  brfss_smart2010 %>%
  filter(year == "2002") %>%
  group_by(state) %>%
  summarize(count_location = n()) %>%
  filter(count_location >= 7)
```

There are `r length(dplyr::pull(more_than_7_2002, state))` states:

`r dplyr::pull(more_than_7_2002, state)`

What about in 2010?
```{r 2010, message=FALSE}
more_than_7_2010 <-
  brfss_smart2010 %>%
  filter(year == "2010") %>%
  group_by(state) %>%
  summarize(count_location = n()) %>%
  filter(count_location >= 7)
```

There are `r length(dplyr::pull(more_than_7_2010, state))` states:

`r dplyr::pull(more_than_7_2010, state)`

2. Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state

```{r spaghetti, message=FALSE}
brfss_smart2010_spaghetti <-
  brfss_smart2010 %>%
  filter(response == "Excellent") %>%
  group_by(year, state) %>%
  summarise(mean_data_value = mean(data_value))

brfss_smart2010_spaghetti
```

Plot:
```{r}
brfss_smart2010_spaghetti %>%
  ggplot(aes(x = year, y = mean_data_value)) +
  geom_line(aes(color = state)) +
  labs(
    title = "Spaghetti Plot for Mean Data Value over Time",
    x = "Year",
    y = " Mean Data Value across Locations within a State",
    caption = "Data from the BRFSS SMART dataset"
  )
```

* Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
brfss_smart2010 %>%
  filter(year %in% c("2006", "2010"),
         state == "NY") %>%
  group_by(response) %>%
  ggplot(aes(x = data_value, fill = response)) +
  geom_density(alpha = .4, adjust = .5, color = "blue") +
  labs(
    title = "Two-Panel Plot for Responses in NY State in 2002 and 2010 Year",
    x = "Data Value",
    y = "Density",
    caption = "Data from the BRFSS SMART dataset"
  ) +
  facet_grid(year ~ response)
```






