---
title: "HW 3"
author: Shihui Zhu sz3029
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set()
```

```{r package, message=FALSE}
# This chunk loads all the packages used in this homework
library(tidyverse)
library(p8105.datasets)
```

# Problem 1

## Load instacart dataset
```{r}
data("instacart")
```

## Description of the dataset

The instacart dataset contains `r nrow(instacart)` observations of `r ncol(instacart)` variables. The variables are: ``r names(instacart)``. Some key variables are: 

* ``r names(instacart)[1]``: order identifier, contains `r length(unique(dplyr::pull(instacart, order_id)))` individual orders

* ``r names(instacart)[2]``: product identifier, contains `r length(unique(dplyr::pull(instacart, product_id)))` individual products

* ``r names(instacart)[4]``: 1 if the product has been ordered by this user in the past, 0 otherwise

* ``r names(instacart)[5]``: customer identifier, contains `r length(unique(dplyr::pull(instacart, user_id)))` individual customers

* ``r names(instacart)[7]``: the order sequence number for this user

* ``r names(instacart)[8]``: the day of the week on which the order was placed

* ``r names(instacart)[12]``: aisle identifier, contains `r length(unique(dplyr::pull(instacart, aisle_id)))` aisles

* ``r names(instacart)[13]``: department identifier, contains `r length(unique(dplyr::pull(instacart, department_id)))` departments

`r instacart[1,] %>% knitr::kable()`

An example of the above observation in the dataset: A customer with id 112108 ordered "Bulgarian Yogurt" (product id 49302) from department "dairy eggs," aisle "yogurt". This was his/her 4th order at instacart, and the product was a re-ordered product for him/her, as well as the first order added to cart. The order was placed at Thursday 10am, and it was 9 days after his/her last order. 

## Questions

* How many aisles are there, and which aisles are the most items ordered from?
```{r}
length(unique(dplyr::pull(instacart, aisle_id)))
```

There are 134 aisles.

```{r}
instacart %>%
  group_by(aisle_id) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  slice(1:1)
```

The aisle with id 83, "fresh vegetables," is which the most items ordered from, and it is ordered 150609 times.

* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>%
  group_by(aisle_id) %>%
  filter(n() > 10000) %>%
  arrange(aisle_id) %>%
  ggplot(aes(x = aisle_id)) +
  geom_histogram(binwidth = 2) +
  labs(
    title = "Histogram for number of items ordered in each aisle (> 10000 items ordered)",
    x = "Aisle ID",
    y = "Number of items ordered")
```

* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>%
  filter(
    aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")
  ) %>%
  group_by(aisle, product_name) %>%
  summarise(count = n()) %>%
  mutate(rank =  dense_rank(desc(count))) %>%
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>%
  knitr::kable(col.names = c("Aisle", "Product Name", "Count", "Rank"))
```

* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers 

```{r}
#week_day <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")
instacart %>%
  filter(
    product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")
  ) %>%
  mutate(order_dow = lubridate::wday(order_dow + 1, label = TRUE)) %>%
  group_by(product_name, order_dow) %>%
  summarise(mean = mean(order_hour_of_day)) %>%
  knitr::kable(digits = 1, col.names = c("Product", "Week Day", "Mean hour of the day"))
```

# Problem 2

## Load BRFSS dataset
```{r}
data("brfss_smart2010")

df_brfss_smart2010_clean <- 
  brfss_smart2010 %>% 
  janitor::clean_names()
```







